### 2020-04-22 paper_results.Rmd
initialize

---
title: "Untitled"
author: "Darren L Dahly"
date: "6 July 2017"
output:
  word_document:  
    reference_docx: "style.1.docx"
---

```{r setup, include = FALSE}

  knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, 
                        fig.width  = 6 * 1.67, fig.height = 6)

  load("data/data.RData") # Created with scripts/data.R from "source" data
  source("scripts/functions.R")

  library(tidyverse)
  library(mixor)
  library(sjPlot)
  library(viridis)

```

# Descriptives

The dataset includes `r nrow patients` patients nested in `r length(levels(patients$ward))` wards.

*Table 1. Patient level characteristics

```{r table_1_variables}

# See table 1 from Corina's draft, p10

  table_1_df <- select(patients, 
    age, gender, diag, comorbid, meds, cfs, must_screen, must_score, 
    dementia:paralysis, restrict_diet, modified_diet, diet_ref, diet_plan
    )

```

```{r table_1}

  data_frame(
    Variable = name.1(table_1_df),
    Missing   = n.miss(table_1_df),
    "Mean SD" = summary.1(table_1_df),
    "Median [IQR]" = med.iqr(table_1_df),
    "(Min, Max)" = min.max(table_1_df)
    ) %>% 
  knitr::kable()

```

*Table 2. Meal-level characteristics

# Plots

*Figure X. Patient level distribution of meal intake values by day and meal. 
```{r}

  ggplot(meals, aes(y = ptid, x = which_meal,
                    fill = factor(mealintake))) +
    geom_tile() +
    scale_fill_viridis("Meal Intake", discrete = TRUE,
                       guide = guide_legend(reverse = TRUE)) +
    xlab("") + ylab("Patient ID") +
    theme(legend.position = "bottom",
          panel.background = element_blank(),
          axis.text.y = element_text(size = 6))

  ggsave("plots/intakes.png", height = 19.05, width = 33.86,
         units = "cm", scale = 0.7)

```

Figure X. Distribution of meal intake values
```{r}

  ggplot(meals, aes(x = mealintake)) + geom_histogram()

  ggplot(meals, aes(x = mealintake)) + geom_histogram() +
    facet_wrap(~meal, ncol = 1)

```

# Models

```{r}

  meals <- arrange(meals, ptid, day, meal)

  basic <- mixor(mealintake ~ meal + day, data = meals,
                 id = ptid, link = "probit")

  summary(basic)

```

```{r code_book}

  print(summarytools::dfSummary(data), method = "render")

```

```{r sysinfo}

  DescTools::SysInfo()

```